\documentclass[12pt]{article}
\usepackage{myfullpage}
\usepackage{fancyhdr}
\usepackage{paralist}
\usepackage{amsmath}
\usepackage{url}
\usepackage{color}
\usepackage{graphicx}
\usepackage{hyperref}

\usepackage{times}
\usepackage{sectsty}
\usepackage{xcolor}
\usepackage{esvect}
\usepackage{float}

\sectionfont{\normalsize\bf}
\subsectionfont{\normalsize\bf}
\usepackage{listings}
\lstset{
  language=bash,
  basicstyle=\ttfamily
}


\input defs.tex

\pagestyle{fancyplain}
\rhead{Instructor: Mo Chen}
\lhead{CMPT 726: Assignment 2}
%\renewcommand{\headrulewidth}{0pt}

\begin{document}

%\thispagestyle{empty}

\subsection*{Assignment 2}
\Heading{Due November 03, 2025 at 11:59pm}


{\bf This assignment is to be done individually.}

\rule{\textwidth}{1mm}

{\bf Important Note:} The university policy on academic dishonesty
(cheating) will be taken very seriously in this course.  You may not
provide or use any solution, in whole or in part, to or by another
student.

You are encouraged to discuss the concepts involved in the questions
with other students.  If you are in doubt as to what constitutes
acceptable discussion, please ask!  Further, please take advantage of
office hours offered by the instructor and the TA if you are having
difficulties with this assignment.

{\bf DO NOT}:

\begin{asparaitem}
\item Give/receive code or proofs to/from other students
\item Use Google to find solutions for assignment
\end{asparaitem}


{\bf DO}:

\begin{asparaitem}
\item Meet with other students to discuss assignment (it is best not to take any notes during such meetings, and to re-work assignment on your own)
\item Use online resources (e.g.~Wikipedia) to understand the concepts needed to solve the assignment.
\end{asparaitem}

\rule{\textwidth}{1mm}

\section*{Submitting Your Assignment}

The assignment must be submitted online on Canvas.
You must submit a report in {\bf PDF format}. You may typeset your assignment in LaTeX or Word, or submit neatly handwritten and scanned solutions. We will not be able to give credit to solutions that are not legible.

\rule{\textwidth}{1mm}

\newpage

\input{questions/q1}
\input{questions/q2}
\input{questions/q3}

\end{document}
